---
title: "CRO_MYBs"
author: "Chenxin Li"
date: "2024-01-04"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Package 
```{r}
library(tidyverse)
library(readxl)

library(RColorBrewer)
library(viridis)
library(rcartocolor)
library(svglite)

library(patchwork)

library(ggtree)
library(ggtreeExtra)
library(treeio)
library(ape)

library(Seurat)
```
# Data 
## Bulk RNA-seq
```{r}
sra_metadata <- read_csv("../../CRO_petal_bulk_RNA/Data/metadata_SRA.csv")
head(sra_metadata, 10)
```

```{r}
bulk_RNAseq_data <- read_csv("../../CRO_petal_bulk_RNA/Results/R_output/bulk_exp_table_long.csv")
head(bulk_RNAseq_data)
```
```{r}
JA_libraries <- bulk_RNAseq_data %>% 
  filter(`Sample ID` %in% c("Shoot Ctrl", "Shoot MeJA 6h", 
                            "Suspension Ctrl", "Sus. MeJA 24hr"))

head(JA_libraries)
```

## Single cell RNA-seq 
```{r}
CRO_RNA <- readRDS("../../CRO_multiome/Results/R_output/CRO_RNA.Rds")
CRO_RNA
```
### Cell type color strip 
```{r}
CRO_RNA@meta.data
```

## Tree object
```{r}
CRO_MYB_tree <- read.tree("../Results/CRO_MYB_annotator/RESULTS/Catharanthus_roseus05FastTree_tree.tre")
```

```{r}
Tip_table <- data.frame(
  label = CRO_MYB_tree$tip.label
) %>% 
  mutate(source = case_when(
    str_detect(label, "CRO") ~ "Cath.",
    T ~ "MYB_annotator"
  )) %>% 
  mutate(label_2 = case_when(
    str_detect(label, "Arabidopsis") ~ str_extract(label, "MYB\\d+"),
    str_detect(label, "CRO") ~ str_remove_all(label, "CRO_|CRO.|_\\d+"),
    T ~ ""
  )) %>% 
  mutate(label_3 = case_when(
    str_detect(label, "Arabidopsis") ~ label_2,
    str_detect(label, "CRO") ~ str_remove_all(label_2, "\\.\\d+"),
    T ~ " "
  )) %>% 
  mutate(label_4 = case_when(
    str_detect(label, "05G006800|04G033370|07G002170") ~ label_3,
    T ~ ""
  )) %>% 
  mutate(species = case_when(
    str_detect(label, "Arabidopsis") ~ "A. thaliana",
    str_detect(label, "^CRO") ~ "C. roseus",
    T ~ "other"
  )) %>% 
  mutate(species2 = case_when(
    str_detect(label, "Arabidopsis") ~ "A. thaliana",
    str_detect(label, "^CRO") ~ "C. roseus",
    str_detect(label, "-lycopersicum-") ~ "S. lycopersicum",
    str_detect(label, "Solanum-tuberosum-") ~ "S. tuberosum",
    T ~ "other"
  )) %>% 
  mutate(label_5 = case_when(
    str_detect(label, "lycoper") ~ str_extract(label, "MYB\\d+"),
    T ~ label_3
  )) %>% 
  mutate(index = 1:n()) 

head(Tip_table)
```

```{r}
CRO_MYB_tree_nice <- full_join(
  CRO_MYB_tree, Tip_table, by = "label"
)
```

# Make tree 
```{r}
tree_all <- CRO_MYB_tree_nice %>% 
  ggtree(
    layout = "circular",
    branch.length = "none"
  ) +
  #geom_tiplab(aes(label = label_3), size = 1.5) +
  geom_tippoint(aes(color = species2))

tree_all
```

## Tree with just Arabidopsis and Cath. 
```{r}
Ara_CRO_nodes <- Tip_table %>% 
  filter(species %in% c("A. thaliana", "C. roseus"))

subset_tree <- treeio::drop.tip(
  CRO_MYB_tree_nice, 
  CRO_MYB_tree$tip.label[!(CRO_MYB_tree$tip.label %in% Ara_CRO_nodes$label)]
                        )
```

```{r}
tree_at_cro <- subset_tree %>% 
  ggtree(
    layout = "circular",
    branch.length = "none"
  ) +
  geom_tippoint(aes(color = species)) +
  geom_tiplab(aes(label = label_3), size = 2.25) +
  scale_color_manual(values = brewer.pal(8, "Set2")[c(1,4)]) +
  theme(
    legend.position = c(0.1, 0.1)
  )

tree_at_cro

ggsave("../Results/R_output/AT_CRO_MYB_tree.svg", height = 8, width = 8)
ggsave("../Results/R_output/AT_CRO_MYB_tree.pdf", height = 8, width = 8)
ggsave("../Results/R_output/AT_CRO_MYB_tree.png", height = 8, width = 8)
```
```{r}
subset_tree %>% 
  ggtree(
    layout = "circular",
    branch.length = "none"
  ) +
  geom_tippoint(aes(color = species)) +
  #geom_tiplab(aes(label = label_3), size = 2.25) +
  scale_color_manual(values = brewer.pal(8, "Set2")[c(1,4)])+
  theme(legend.position = "none") 

ggsave("../Results/R_output/AT_CRO_MYB_tree_2.svg", height = 3, width = 3)
ggsave("../Results/R_output/AT_CRO_MYB_tree_2.png", height = 3, width = 3)
```

## Tree with Arabidopsis, Catharanthus, tomato and potato 
```{r}
Ara_CRO_Solan_nodes <- Tip_table %>% 
  filter(species2 %in% c("A. thaliana", "C. roseus", "S. tuberosum", "S. lycopersicum"))

subset_tree2 <- treeio::drop.tip(
  CRO_MYB_tree_nice, 
  CRO_MYB_tree$tip.label[!(CRO_MYB_tree$tip.label %in% Ara_CRO_Solan_nodes$label)]
                        )

write_excel_csv(Ara_CRO_Solan_nodes, "../Results/R_output/Ara_CRO_Solan_nodes.csv")
```

```{r}
subset_tree2 %>% 
  ggtree(
    layout = "circular",
    branch.length = "none"
  ) +
  geom_tippoint(aes(color = species2)) +
  geom_tiplab(aes(label = label_5), size = 2.25) +
  scale_color_manual(values = c(
    brewer.pal(8, "Set2")[c(1,4)],
    "tomato3", "gold"
  )
    ) +
  labs(color = "species") +
  theme(
   legend.position = c(0.1, 0.1) 
  ) 

ggsave("../Results/R_output/AT_CRO_Solan_MYB_tree.svg", height = 16, width = 16)
ggsave("../Results/R_output/AT_CRO_Solan_MYB_tree.pdf", height = 16, width = 16)
ggsave("../Results/R_output/AT_CRO_Solan_MYB_tree.png", height = 16, width = 16)
```

## Check Li's MYB of interest 
### 05G006800.1 "MYB1"
```{r}
Tip_table %>% 
  filter(label_2 %in% c("MYB000",
                        "MYB023",
                        "MYB066")) 
```

* MYB066 - werewolf - root hair and epidermis
* MYB023 - trichome 
* MYB000 - GL1 - trichome 

```{r}
mrca_node <- getMRCA(CRO_MYB_tree, tip = c("Arabidopsis-thaliana-At3g27920-At2R-MYB000", "CRO-05G006800.1")) 
descendant_tips <- CRO_MYB_tree$tip.label[offspring(CRO_MYB_tree, mrca_node)]

subset_tree_05G006800 <- treeio::drop.tip(
  CRO_MYB_tree_nice, 
  CRO_MYB_tree$tip.label[!(CRO_MYB_tree$tip.label %in% descendant_tips)]
                        )

descendant_tips
```
 
```{r}
subset_tree_05G006800 %>% 
  ggtree(
    layout = "circular",
    branch.length = "none"
  ) +
  geom_tippoint(aes(color = species)) +
  geom_tiplab(aes(label = label), size = 2.25) 
```
### 04G033370
```{r}
Tip_table %>% 
  filter(label_2 %in% c("MYB008",
                        "MYB006")) 
```

* MYB008
* MYB006 

```{r}
mrca_node2 <- getMRCA(CRO_MYB_tree, tip = c("Arabidopsis-thaliana-At4g09460-At2R-MYB006", "CRO-04G033370.1")) 
descendant_tips2 <- CRO_MYB_tree$tip.label[offspring(CRO_MYB_tree, mrca_node2)]

subset_tree_04G033370 <- treeio::drop.tip(
  CRO_MYB_tree_nice, 
  CRO_MYB_tree$tip.label[!(CRO_MYB_tree$tip.label %in% descendant_tips2)]
                        )

descendant_tips2
```
```{r}
subset_tree_04G033370 %>% 
  ggtree(
    layout = "circular",
    branch.length = "none"
  ) +
  geom_tippoint(aes(color = species)) +
  geom_tiplab(aes(label = label), size = 2.25) 
```

### 07G002170
```{r}
Tip_table %>% 
  filter(label_2 %in% c("MYB123")) 
```
* MYB123 TT2 proanthocyanidin accumulation of developing seed

```{r}
mrca_node3 <- getMRCA(CRO_MYB_tree, tip = c("Arabidopsis-thaliana-At5g35550-At2R-MYB123", "CRO-07G002170.1")) 
descendant_tips3 <- CRO_MYB_tree$tip.label[offspring(CRO_MYB_tree, mrca_node3)]

subset_tree_07G002170 <- treeio::drop.tip(
  CRO_MYB_tree_nice, 
  CRO_MYB_tree$tip.label[!(CRO_MYB_tree$tip.label %in% descendant_tips3)]
                        )

descendant_tips3
```

```{r}
subset_tree_07G002170 %>% 
  ggtree(
    layout = "circular",
    branch.length = "none"
  ) +
  geom_tippoint(aes(color = species)) +
  geom_tiplab(aes(label = label), size = 2.25) 
```

# Add JA treatment logFC
```{r}
tree_at_cro2 <- subset_tree %>% 
  ggtree(
    #layout = "circular",
    #branch.length = "none"
  ) +
  geom_tippoint(aes(color = species), size = 1) +
  #geom_tiplab(aes(label = label_3), size = 2.25) +
  #theme(legend.position = c(0.8, 0.2)) +
  scale_color_manual(values = brewer.pal(8, "Set2")[c(1,4)])

tree_at_cro2
```

```{r}
JA_libraries_wide <- JA_libraries %>% 
  mutate(tag = str_remove(target_id, "CRO_")) %>% 
  mutate(tag = str_remove(tag, "\\.\\d+")) %>%
  mutate(gene_ID = str_replace(target_id, "_", "-")) %>% 
  select(gene_ID, tag, tpm, `Sample ID`) %>% 
  #filter(label_3 %in% Tip_table$label_3) %>% 
  pivot_wider(names_from = `Sample ID`, values_from = tpm) %>% 
  mutate(shoot_Log2FC = log2((`Shoot MeJA 6h` + 1) / (`Shoot Ctrl` + 1))) %>% 
  mutate(sus_Log2FC = log2((`Sus. MeJA 24hr` + 1) / (`Suspension Ctrl` + 1))) 

head(JA_libraries_wide)
```
```{r}
JA_libraries_wide %>% 
  filter(tag %in% Tip_table$label_3) %>% 
  #filter(shoot_Log2FC > 1) %>% 
  #filter(sus_Log2FC > 1) %>% 
  ggplot(aes(x = shoot_Log2FC, y = sus_Log2FC)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_classic() +
  coord_fixed()
```

```{r}
TF_candidate_wide <- JA_libraries_wide %>% 
  filter(tag %in% c("03G000120", 
                          "04G033370", "05G006800", "07G002170")) %>%
   mutate(tag2 = case_when(
    tag == "03G000120" ~ "IDW1",
    tag == "04G033370" ~ "IDM2",
    tag == "05G006800" ~ "IDM1",
    tag == "07G002170" ~ "IDM3"
  )) %>% 
  mutate(tag2 = factor(tag2, levels = c(
     "IDW1", "IDM3", "IDM2", "IDM1"
  ))) 

TF_candidate_wide
```

## JA response of idioblast TF candidates 
```{r}
TF_JA_lfc_plot <- TF_candidate_wide %>% 
  ggplot(aes(x = tag2, y = "24 hr MeJA / ctrl")) +
  geom_tile(aes(fill = sus_Log2FC)) +
  scale_fill_gradientn(colors = rev(brewer.pal(11, "RdBu")),
                       limits = c(-5.5, 5.5),
                       breaks = c(-5, 0, 5)) +
  labs(x = "TFs",
       y = "",
       fill = "log2FC") + 
  theme_classic() +
  theme(axis.text = element_text(color = "black")) 
 

TF_JA_tpm_plot <- TF_candidate_wide %>% 
  dplyr::select(tag2, `Suspension Ctrl`, `Sus. MeJA 24hr`) %>% 
  pivot_longer(cols = !tag2, names_to = "treatment", values_to = "tpm") %>% 
  mutate(log2TPM = log2(tpm + 1)) %>% 
  ggplot(aes(x = tag2, y = treatment)) +
  geom_tile(aes(fill = log2TPM)) +
  scale_fill_gradientn(colors = brewer.pal(9, "YlGnBu"),
                       #limits = c(-5.5, 5.5),
                       breaks = c(1, 3, 5)
                       ) +
  labs(x = "TFs",
       y = "treatment",
       fill = "log2(TPM + 1)") + 
  theme_classic() +
  theme(axis.text = element_text(color = "black")) 

wrap_plots(TF_JA_tpm_plot, TF_JA_lfc_plot,
           nrow = 2, heights = c(1, 0.6),
           guides = "collect") &
  theme(legend.key.height = unit(0.8, "lines"),
        legend.key.width = unit(0.7, "lines"))

ggsave("../Results/R_output/IDW_IDM_JA.svg", height = 3, width = 6)
ggsave("../Results/R_output/IDW_IDM_JA.png", height = 3, width = 6)
```

```{r}
TF_JA_tpm_plot

ggsave("../Results/R_output/IDW_IDM_JA-1.svg", height = 2, width = 5.5)
ggsave("../Results/R_output/IDW_IDM_JA-1.png", height = 2, width = 5.5)
```


```{r}
JA_LFC <- JA_libraries_wide %>% 
  filter(tag %in% Tip_table$label_3) %>% 
  dplyr::select(tag, gene_ID, shoot_Log2FC, sus_Log2FC) %>% 
  pivot_longer(names_to = "experiment", values_to = "log2FC", cols = contains("Log2FC")) %>% 
  mutate(Treatment = case_when(
    str_detect(experiment, "shoot") ~ "Shoot MeJA 6hr",
    str_detect(experiment, "sus") ~ "Sus. MeJA 6hr",
  )) %>% 
  mutate(log2FC_clipped = case_when(
    log2FC > 2 ~ 2,
    log2FC < -1 ~ -1,
    T ~ log2FC
  )) %>% 
  mutate(status = case_when(
    log2FC > 1 ~ "JA upregulated",
    T ~ "other"
  ))
  

head(JA_LFC)
```

```{r}
summary(JA_LFC$log2FC)
quantile(JA_LFC$log2FC, c(0.025, 0.05, 0.1, 0.9, 0.95, 0.975))
```

## Assign JA-MYBs 
```{r}
JA_MYBs <- JA_libraries_wide %>% 
  filter(tag %in% Tip_table$label_3) %>% 
  mutate(status = case_when(
    shoot_Log2FC > 1 |  
      sus_Log2FC > 1 ~ "JA up MYB",
    T ~ "other"
  )) %>% 
  select(gene_ID, shoot_Log2FC, sus_Log2FC,status)  

JA_MYBs
```
 


```{r}
JA_MYBs %>% 
  group_by(status) %>% 
  dplyr::count()

JA_MYBs %>% 
  filter(status == "JA up MYB") %>% 
  arrange(gene_ID)
```

18 JA MYBs 
The 3 MYBs that I pulled out from idioblast are all JA MYB (up in the suspension culture experiment). 

# Add single cell data 
```{r}
# JA_MYBs$gene_ID
MYB_dotplot <- DotPlot(CRO_RNA, features = JA_MYBs$gene_ID, group.by = "cell_type")$data  %>% 
  mutate(id2 = case_when(
    id == "Mesophyll" ~ "M",
    id == "Epidermis" ~ "E",
    id == "Guard cells" ~ "GC",
    id == "Vasculature" ~ "V",
    id == "IPAP" ~ "IP",
    id == "Idioblast" ~ "Id",
    T ~ "Un"
  )) %>% 
  mutate(id2 = factor(id2, levels = c(
    "IP", "E", "Id", "M", "GC", "V", "Un" 
  )))

MYB_dotplot
```

## Assign IPAP, Epi, and Id MYBs 
 

```{r}
MYB_classification <- MYB_dotplot %>% 
  mutate(gene_ID = features.plot) %>% 
  group_by(gene_ID) %>% 
  slice_max(n = 1, order_by = avg.exp.scaled) %>% 
  mutate(peak_exp = id2) %>% 
  select(gene_ID, peak_exp, avg.exp.scaled) %>% 
  full_join(JA_MYBs, by = "gene_ID") %>% 
  arrange(gene_ID)

head(MYB_classification)
```
```{r}
Interesting_MYBs <- MYB_classification %>% 
  filter(status == "JA up MYB" | 
           peak_exp %in% c("Id", "IP", "E")) %>% 
  arrange(peak_exp)

Interesting_MYBs
write_excel_csv(Interesting_MYBs, "../Results/R_output/Interesting_MYBs.csv")
```
```{r}
Interesting_MYBs %>% 
  group_by(peak_exp) %>% 
  dplyr::count()

MYB_classification %>% 
  group_by(peak_exp) %>% 
  dplyr::count() %>% 
  dplyr::rename(total = n)
```
```{r}
JA_MYBs_at_cell_type <- MYB_classification %>% 
  group_by(peak_exp, status) %>% 
  dplyr::count() %>% 
  ungroup() %>% 
  mutate(
    peak_exp2 = case_when(
      is.na(peak_exp) ~ "ND",
      T ~ as.character(peak_exp)
    )
  ) %>% 
  inner_join(
    MYB_classification %>% 
  group_by(peak_exp) %>% 
  dplyr::count() %>% 
  dplyr::rename(total = n), 
  by = "peak_exp"
  ) %>% 
  mutate(percent = n / total * 100) %>% 
  mutate(peak_exp2 = factor(peak_exp2, levels = c(
   "Id", "E", "V", "Un","GC",  "IP", "ND", "M"
  ))) 
  
JA_MYBs_at_cell_type %>% 
  ggplot(aes(x = peak_exp2, y = percent)) +
  geom_bar(stat = "identity", aes(fill = status)) +
  scale_fill_manual(values  = c("tomato1", "grey80")) +
  labs(x = "Cell type of highest expression",
       y = "% MYBs") +
  theme_classic() +
  theme(
    legend.position = "top"
  )

ggsave("../Results/R_output/JA_MYBs_at_cell_type.svg", height = 2.5, width = 3)
ggsave("../Results/R_output/JA_MYBs_at_cell_type.png", height = 2.5, width = 3)
```
```{r}
JA_MYBs_at_cell_type
```

```{r}
Interesting_MYBs %>% 
  filter(status == "JA up MYB" |
           peak_exp == "Id")
```

# Wrap plots together 
```{r}
tree_at_cro3 <- tree_at_cro2 +
  labs(color = "Species") +
  geom_fruit(data = JA_MYBs,
             geom = geom_point,
             shape = 21, color = "white",
             size = 1,
             aes(y = gene_ID, x = 0, fill = status),
             pwidth = 0.01) +
  scale_fill_manual(values  = c("tomato1", "grey30")) +
  labs(fill = "Response to JA") +
  ggnewscale::new_scale_fill() +
  geom_fruit(
     data = MYB_dotplot,
     geom = geom_tile,
     color = "grey90",
     aes(y = features.plot, x = id2, fill = avg.exp.scaled),
     axis.params = list(axis = "x", text.size = 2, vjust = -0.15)
   ) +
  scale_fill_gradientn(colors = viridis(n = 10, option = "A", begin = 0, end = 0.95)) +
  labs(fill = "Avg. Exp.\nscRNA-seq") +
  theme(legend.position = "left")


tree_at_cro3

ggsave("../Results/R_output/AT_CRO_MYB_tree_extra.svg", height = 8, width = 6)
ggsave("../Results/R_output/AT_CRO_MYB_tree_extra.pdf", height = 8, width = 6)
ggsave("../Results/R_output/AT_CRO_MYB_tree_extra.png", height = 8, width = 6)
```

## A different version with idioblast MYB1 labeled 
```{r}
tree_at_cro2 +
  geom_tiplab(aes(label = label_4), size = 3, hjust = -0.5, vjust = 0.5) +
  labs(color = "Species") +
  geom_fruit(data = JA_MYBs,
             geom = geom_point,
             shape = 21, color = "white",
             size = 1,
             aes(y = gene_ID, x = 0, fill = status),
             pwidth = 0.01) +
  scale_fill_manual(values  = c("tomato1", "grey30")) +
  labs(fill = "Response to JA") +
  ggnewscale::new_scale_fill() +
  geom_fruit(
     data = MYB_dotplot,
     geom = geom_tile,
     color = "grey90",
     aes(y = features.plot, x = id2, fill = avg.exp.scaled),
     axis.params = list(axis = "x", text.size = 2, vjust = -0.15)
   ) +
  scale_fill_gradientn(colors = viridis(n = 10, option = "A", begin = 0, end = 0.95)) +
  scale_y_continuous(breaks = seq(from = 3, to = nrow(Tip_table)-2, by = 2)) +
  labs(fill = "Avg. Exp.\nscRNA-seq") +
  theme(legend.position = "left",
        panel.grid.major.y = element_line(),
        panel.grid.minor.y = element_line())

ggsave("../Results/R_output/AT_CRO_MYB_tree_extra_2.svg", height = 8, width = 6)
ggsave("../Results/R_output/AT_CRO_MYB_tree_extra_2.pdf", height = 8, width = 6)
ggsave("../Results/R_output/AT_CRO_MYB_tree_extra_2.png", height = 8, width = 6)
```

 
# Load bHLH genes  
```{r}
bHLH_list <- read_delim("../Results/CRO_bHLH_annotator/RESULTS/Catharanthus_roseus04b_candidate_2_reference_mapping_file.txt", 
                        delim = "\t")

head(bHLH_list)
```
## JA up bHLHs
```{r}
JA_libraries_wide
```

```{r}
JA_bHLH <- JA_libraries_wide %>% 
  filter(gene_ID %in% bHLH_list$NewMember) %>% 
  mutate(status = case_when(
    shoot_Log2FC > 1 |  
      sus_Log2FC > 1 ~ "JA up bHLH",
    T ~ "other"
  )) %>% 
  select(gene_ID, shoot_Log2FC, sus_Log2FC,status)  

head(JA_bHLH)
```
```{r}
JA_bHLH %>% 
  group_by(status) %>% 
  dplyr::count()
```

* 27 JA up bHLH
* 210 non-JA up bHLH

```{r}
JA_bHLH %>% 
  filter(status == "JA up bHLH") %>% 
  arrange(gene_ID)
```

04G007340.2 is JA up  

## bHLH at single cell  
```{r}
bHLH_heat_data <- DotPlot(CRO_RNA, features = bHLH_list$NewMember)$data

bHLH_peak_exp <- bHLH_heat_data %>% 
  group_by(features.plot) %>% 
  slice_max(n = 1, order_by = avg.exp.scaled) %>% 
  ungroup() %>% 
  select(features.plot, id) %>% 
  dplyr::rename(peak_cluster = id) %>% 
  mutate(order_y = as.numeric(peak_cluster))

head(bHLH_peak_exp)
```
```{r}
bHLH_heat_data %>% 
  left_join(bHLH_peak_exp, by = "features.plot") %>% 
  mutate(features.plot = reorder(features.plot, -order_y)) %>% 
  ggplot(aes(x = id, y = features.plot)) +
  geom_tile(aes(fill = avg.exp.scaled)) +
  scale_fill_gradientn(colors = brewer.pal(9, "YlGnBu")) +
  labs(y = "bHLH genes",
       x = "cell cluters",
       fill = "Avg. Exp") +
  theme_classic() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text = element_text(color = "black"),
        legend.position = "top")


```
```{r}
idioblast_bHLH <- bHLH_peak_exp %>% 
  filter(peak_cluster == 8)
```

```{r}
DotPlot(CRO_RNA, features = rev(idioblast_bHLH$features.plot)) +
  coord_flip()
```
A potential bHLHs that interacts with idioblast MYBs: 04G007340.2 

```{r}
bHLH_list %>% 
  filter(NewMember %in% idioblast_bHLH$features.plot) %>% 
  select(-EdgeDistance, -PatristicDistance)
```

CRO-04G007340.2	seems to be the interesting one 

```{r}
bHLH_list %>% 
  inner_join(JA_bHLH, by = c("NewMember"="gene_ID")) %>% 
  filter(status == "JA up bHLH") %>% 
  select(-EdgeDistance, -PatristicDistance)
```

```{r}
bHLH_list %>% 
  inner_join(JA_bHLH, by = c("NewMember"="gene_ID")) %>% 
  filter(status == "JA up bHLH") %>% 
  filter(NewMember %in% idioblast_bHLH$features.plot) %>% 
  select(-EdgeDistance, -PatristicDistance)
```

3 bHLHs that are both JA up and idioblast specific. 
