---
title: "Multi_GEX_starsolo"
author: "Chenxin Li"
date: "2023-03-31"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# packages 
```{r}
library(tidyverse)
library(Seurat)
library(readxl)
library(RColorBrewer)
library(rcartocolor)
library(viridis)
library(patchwork)
```

# read data 
```{r}
AZ_raw <- Read10X("../Results/22453Bue_CRO-AZ_S3_L002_GeneFull/raw")
BB_raw <- Read10X("../Results/22453Bue_CRO-BB_S4_L002_GeneFull/raw") 
BD_raw <- Read10X("../Results/22453Bue_CRO-BD_S5_L002_GeneFull/raw") 
```

# objects 
```{r}
AZ <- CreateSeuratObject(counts = AZ_raw, min.cells = 3, min.features = 200,
                         project = "AZ")

BB <- CreateSeuratObject(counts = BB_raw, min.cells = 3, min.features = 200,
                         project = "BB")

BD <- CreateSeuratObject(counts = BD_raw, min.cells = 3, min.features = 200,
                         project = "BD")
```

```{r}
AZ
BB
BD
```

# Filtering 
```{r}
blank <- data.frame(
  x = 1:10,
  y = 1:10
) %>% 
  ggplot(aes(x = x, y = y)) +
  theme_void()
```

## AZ 
```{r}
AZ_scatter <- AZ@meta.data %>% 
  ggplot(aes(x = nCount_RNA, y = nFeature_RNA)) +
  geom_point(alpha = 0.5) +
  geom_vline(xintercept = 300, color = "red2") +
  geom_vline(xintercept = 5000, color = "red2") +
  geom_hline(yintercept = 300, color = "blue1") +
  geom_hline(yintercept = 3000, color = "blue1") +
  scale_x_log10() +
  scale_y_log10() +
  theme_classic()

AZ_hist1 <- AZ@meta.data %>% 
  ggplot(aes(x = nCount_RNA)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 300, color = "red2") +
  geom_vline(xintercept = 5000, color = "red2") +
  scale_x_log10() +
  theme_classic()

AZ_hist2 <- AZ@meta.data %>% 
  ggplot(aes(x = nFeature_RNA)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 300, color = "blue1") +
  geom_vline(xintercept = 3000, color = "blue1") +
  scale_x_log10() +
  theme_classic() +
  coord_flip()

AZ_QC <- wrap_plots(
  AZ_hist1 + 
    labs(tag = "A"), blank,
  AZ_scatter +
    labs(title = "rep1"), AZ_hist2,
  nrow = 2, ncol = 2, 
  widths = c(1, 0.25), 
  heights = c(0.28, 1)
)

AZ_QC
ggsave("../Results/R_output/AZ_filtering.svg", height = 4, width = 5, bg = "white")
ggsave("../Results/R_output/AZ_filtering.png", height = 4, width = 5, bg = "white")
```
```{r}
AZ_sub <- subset(AZ, subset = nFeature_RNA > 300 &
                   nFeature_RNA < 3000 &
                   nCount_RNA < 5000 &
                   nCount_RNA > 300)

AZ_sub
```

## BB
```{r}
BB_scatter <- BB@meta.data %>% 
  ggplot(aes(x = nCount_RNA, y = nFeature_RNA)) +
  geom_point(alpha = 0.5) +
  geom_vline(xintercept = 300, color = "red2") +
  geom_vline(xintercept = 25000, color = "red2") +
  geom_hline(yintercept = 300, color = "blue1") +
  geom_hline(yintercept = 10000, color = "blue1") +
  scale_x_log10() +
  scale_y_log10() +
  theme_classic()

BB_hist1 <- BB@meta.data %>% 
  ggplot(aes(x = nCount_RNA)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 300, color = "red2") +
  geom_vline(xintercept = 25000, color = "red2") +
  scale_x_log10() +
  theme_classic()

BB_hist2 <- BB@meta.data %>% 
  ggplot(aes(x = nFeature_RNA)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 300, color = "blue1") +
  geom_vline(xintercept = 10000, color = "blue1") +
  scale_x_log10() +
  theme_classic() +
  coord_flip()

BB_QC <- wrap_plots(
  BB_hist1 +
    labs(tag = "B"), blank,
  BB_scatter +
    labs(title = "rep2"), BB_hist2,
  nrow = 2, ncol = 2, 
  widths = c(1, 0.25), 
  heights = c(0.27, 1)
)

BB_QC
ggsave("../Results/R_output/BB_filtering.svg", height = 4, width = 5, bg = "white")
ggsave("../Results/R_output/BB_filtering.png", height = 4, width = 5, bg = "white")
```

```{r}
BB_sub <- subset(BB, subset = nFeature_RNA > 300 &
                   nFeature_RNA < 10000 &
                   nCount_RNA < 25000 &
                   nCount_RNA > 300)

BB_sub
```

## BD
```{r}
BD_scatter <- BD@meta.data %>% 
  ggplot(aes(x = nCount_RNA, y = nFeature_RNA)) +
  geom_point(alpha = 0.5) +
  geom_vline(xintercept = 250, color = "red2") +
  geom_vline(xintercept = 10000, color = "red2") +
  geom_hline(yintercept = 250, color = "blue1") +
  geom_hline(yintercept = 5000, color = "blue1") +
  scale_x_log10() +
  scale_y_log10() +
  theme_classic()

BD_hist1 <- BD@meta.data %>% 
  ggplot(aes(x = nCount_RNA)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 250, color = "red2") +
  geom_vline(xintercept = 10000, color = "red2") +
  scale_x_log10() +
  theme_classic()

BD_hist2 <- BD@meta.data %>% 
  ggplot(aes(x = nFeature_RNA)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 250, color = "blue1") +
  geom_vline(xintercept = 5000, color = "blue1") +
  scale_x_log10() +
  theme_classic() +
  coord_flip()

BD_QC <- wrap_plots(
  BD_hist1 +
    labs(tag = "C"), blank,
  BD_scatter +
    labs(title = "rep3"), BD_hist2,
  nrow = 2, ncol = 2, 
  widths = c(1, 0.25), 
  heights = c(0.27, 1)
)

BD_QC

ggsave("../Results/R_output/BD_filtering.svg", height = 4, width = 5, bg = "white")
ggsave("../Results/R_output/BD_filtering.png", height = 4, width = 5, bg = "white")
```

```{r}
BD_sub <- subset(BD, subset = nFeature_RNA > 250 &
                   nFeature_RNA < 5000 &
                   nCount_RNA < 10000 &
                   nCount_RNA > 250)

BD_sub
```

## Summaries 
```{r}
summary(AZ_sub$nCount_RNA)
summary(AZ_sub$nFeature_RNA)
```
```{r}
summary(BB_sub$nCount_RNA)
summary(BB_sub$nFeature_RNA)
```

```{r}
summary(BD_sub$nCount_RNA)
summary(BD_sub$nFeature_RNA)
```
# Normalize & variable features 
```{r}
AZ_sub <- NormalizeData(AZ_sub, normalization.method = "LogNormalize", scale.factor = 10000)
BB_sub <- NormalizeData(BB_sub, normalization.method = "LogNormalize", scale.factor = 10000)
BD_sub <- NormalizeData(BD_sub, normalization.method = "LogNormalize", scale.factor = 10000)
```

```{r}
AZ_sub <- FindVariableFeatures(AZ_sub, selection.method = "vst", nfeatures = 3000)
BB_sub <- FindVariableFeatures(BB_sub, selection.method = "vst", nfeatures = 3000)
BD_sub <- FindVariableFeatures(BD_sub, selection.method = "vst", nfeatures = 3000)
```
# Integrate 
```{r}
RNA.list <- list(AZ_sub, BB_sub, BD_sub)

features <- SelectIntegrationFeatures(object.list = RNA.list, nfeatures = 3000)

anchors <- FindIntegrationAnchors(object.list = RNA.list, anchor.features = features) 

CRO_RNA <- IntegrateData(anchorset = anchors)
```

# Scale data 
```{r}
DefaultAssay(CRO_RNA) <- "integrated"
CRO_RNA <- ScaleData(CRO_RNA, verbose = F)

CRO_RNA@meta.data$run <- CRO_RNA@meta.data$orig.ident

CRO_RNA@meta.data %>% 
  group_by(run) %>% 
  dplyr::count()

CRO_RNA
```

# PCA
```{r}
CRO_RNA <- RunPCA(CRO_RNA, verbose = FALSE, npcs = 50) # run PCA

DimPlot(CRO_RNA, reduction = "pca", cols = brewer.pal(8, "Set2")[c(1:3)],
        split.by = "run", group.by = "run") +
theme(
  legend.position = "none"
)
```

# UMAP
```{r}
CRO_RNA <- RunUMAP(CRO_RNA, dims = 1:30, 
                    min.dist = 0.001, repulsion.strength = 1, n.neighbors = 15, spread = 5) 

CRO_RNA <- FindNeighbors(CRO_RNA, reduction = "pca", dims = 1:50)
CRO_RNA <- FindClusters(CRO_RNA, resolution = 0.5)
```

```{r}
UMAP_RNA <- DimPlot(CRO_RNA, reduction = "umap", 
                 label = T, label.size = 5, repel = T) + 
  theme_void() +
  theme(
    text = element_text(size = 14, color = "black"),
    legend.position = "none"
  )

UMAP_RNA
ggsave("../Results/R_output/UMAP_RNA.svg", height = 3, width = 3, bg = "white")
ggsave("../Results/R_output/UMAP_RNA.png", height = 3, width = 3, bg = "white")
```

## Check replicates 
```{r}
UMAP_by_rep <- DimPlot(CRO_RNA, reduction = "umap", label = T, label.size = 5, repel = T, split.by = "run") +
  theme_void() +
  theme(
    text = element_text(size = 14, color = "black"),
    legend.position = "none"
  ) +
  ggtitle("Grouped by replicates\n")  
  
UMAP_by_rep
ggsave("../Results/R_output/UMAP_RNA_by_rep.svg", height = 3, width = 7.5, bg = "white")
ggsave("../Results/R_output/UMAP_RNA_by_rep.png", height = 3, width = 7.5, bg = "white")
```

# Heat map at cluster level (MIA genes)
## Pull annotation
```{r}
repr_gene_models <- read_delim("../Data/cro_v3_anno/cro_v3.gene_models.repr.gff3",
                               delim = "\t", col_names = F)

repr_loci <- repr_gene_models %>% 
  dplyr::filter(X3 == "mRNA") %>% 
  dplyr::select(X1, X4, X5, X7, X9) %>% 
  separate(X9, c("ID", "Name"), sep = ";") %>% 
  dplyr::select(-Name) %>% 
  separate(ID, c("temp", "gene_ID"), sep = "=") %>% 
  dplyr::select(-temp) %>% 
  mutate(LocusID = str_sub(gene_ID, start = 1, end = 13)) %>% 
  mutate(gene_ID = str_replace(gene_ID, "_", "-")) %>% 
  mutate(LocusID = str_replace(LocusID, "_", "-")) %>% 
  dplyr::rename(
    Chr = X1,
    start = X4,
    end = X5,
    strand = X7
  )


head(repr_loci)
```
## Pull MIA genes 
```{r}
MIA_genes_info <- readRDS(file = "../Data/MIA_genes_info.Rds")
leaf_MIA_genes <- MIA_genes_info %>% 
  mutate(tag = case_when(
    str_detect(as.character(gene), "05G028810") ~ "7DLGT",
    str_detect(as.character(gene), "04G032090") ~ "GPPS(LSU)",
    str_detect(as.character(gene), "05G017180") ~ "DPAS", 
    T ~ tag
    )) %>% 
  dplyr::filter(is.na(tag) == F) %>% 
  mutate(order = case_when(
    str_detect(as.character(gene), "04G032090") ~  9,
    str_detect(as.character(gene), "05G028810") ~ 15, 
    str_detect(tag, "THAS2") ~ 40,
    T ~ order
  )) %>% 
  arrange(order) %>%
  mutate(order2 = case_when(
    str_detect(as.character(gene), "04G032090") ~  9,
    str_detect(as.character(gene), "05G028810") ~ 15, 
    str_detect(as.character(gene), "05G017180") ~ 31,
    order >= 31 ~ order + 1,
    T ~ order
  )) %>% 
  arrange(order2) %>% 
  mutate(segment = case_when(
    str_detect(as.character(gene), "04G032090") ~ "MEP",
    str_detect(as.character(gene), "05G028810") ~ "Iridoid", 
    T ~ segment
  )) %>% 
  dplyr::filter(is.na(order2) == F) %>% 
  dplyr::filter(str_detect(as.character(gene), "05G025350") == F) %>% 
  dplyr::filter(str_detect(as.character(gene), "05G028780") == F) %>% 
  dplyr::filter(str_detect(tag, "THAS1|THAS3|THAS4|HYS|ISP|GS2") == F) %>% 
  mutate(tag = case_when(
    str_detect(tag, "SGD$") ~ "SGD1",
    #str_detect(tag, "SGD2") ~ "SDG2",
    str_detect(tag, "GS$") ~ "GS1",
    str_detect(tag, "MATE") ~ "SLTr",
    T ~ tag
  )) %>% 
  mutate(gene = str_replace_all(gene, "_", "-")) %>% 
  left_join(repr_loci %>% 
              dplyr::select(-Chr, -start, -end, -strand), by = c("gene"="LocusID"))

head(leaf_MIA_genes)
write_excel_csv(leaf_MIA_genes, "../Results/R_output/leaf_MIA_genes.csv")
```
```{r}
DefaultAssay(CRO_RNA) <- "RNA"
MIA_at_cluster <- DotPlot(CRO_RNA, features = rev(leaf_MIA_genes$gene_ID))

MIA_heatmap <- MIA_at_cluster$data %>% 
  mutate(internal.tag = row.names(.)) %>% 
  arrange(internal.tag) %>% 
  mutate(gene_ID = case_when(
    is.na(features.plot) ~ str_remove_all(internal.tag, "rna_"),
    T ~ features.plot %>% as.character()
  )) %>% 
  mutate(gene_ID = str_sub(gene_ID, start = 1, end = 15)) %>% 
  inner_join(leaf_MIA_genes, by = "gene_ID") %>%
  mutate(gene = reorder(gene_ID, -order2)) %>%
  mutate(tag2 = case_when(
    str_detect(tag, "DXS|DXR|MCS|GES|G8H|GOR|IO|7DL|LAMT|SLS|TDC|STR&|SDG1|16OM|D4H|DAT") ~ paste0("*", tag),
    T ~ tag
  )) %>% 
  mutate(tag2 = reorder(tag2, -order2)) %>% 
  mutate(id = factor(id, levels = c(
    "0", "1", "3", "7", 
    "2","5", 
    "12",
    "11", "6", "10", "14", "15",
    "13", "8",
    "4", "9"
  ))) %>% 
  ggplot(aes(y = tag2, x = id)) +
  geom_point(aes(size = pct.exp, fill = avg.exp.scaled), shape = 21, color = "white") +
  scale_fill_viridis(option = "A", begin = 0, end = 0.9) +
  guides(size = guide_legend(override.aes = list(fill = "grey40"))) +
  labs(x = "Cluster",
       y = NULL, 
       fill = "Avg.\nExp.",
       size = "% Exp.") +
  theme_classic() +
  theme(text = element_text(size = 14, color = "black"),
        axis.text = element_text(color = "black"),
        axis.text.y.left = element_text(hjust = 0.5),
        legend.key.width = unit(0.8, "lines"))

MIA_heatmap
```

# Pull marker genes 
```{r}
leaf_markers <- read_delim("../Data/leaf_markers.csv", col_types = cols())

leaf_markers <- leaf_markers %>% 
  mutate(gene = str_replace(cro.v3, "_", "-")) %>% 
  left_join(repr_loci, by = c("gene" = "LocusID"))

head(leaf_markers)
```

### CRO specific cells 
```{r}
cro_special <- repr_loci %>% 
  dplyr::filter(LocusID == "CRO-03G004950" | # IO
                  LocusID == "CRO-03G018360" | # 7DLH 
                   LocusID == "CRO-06G013360" | # D4H
                  LocusID == "CRO-02G001090") %>%  # DAT
  cbind(
    Symbol = c("DAT", "IO", "7DLH", "D4H"),
    cell_type = c("idioblast", "IPAP", "IPAP", "idioblast")
  ) %>% 
  dplyr::select(Symbol, cell_type, gene_ID)

cro_special
```

## Cell type assignment heatmap
```{r}
DefaultAssay(CRO_RNA) <- "RNA"

leaf_markers2 <- leaf_markers %>% 
  dplyr::select(Symbol, cell_type, gene_ID) %>% 
  rbind(cro_special) %>% 
  dplyr::filter(str_detect(gene_ID, "007040") == F) %>% 
  dplyr::filter(str_detect(gene_ID, "004679") == F) %>%
  dplyr::filter(str_detect(gene_ID, "023500") == F) %>%
  dplyr::filter(str_detect(gene_ID, "008230") == F) %>%
  dplyr::filter(str_detect(gene_ID, "021030") == F) %>%
  dplyr::filter(str_detect(gene_ID, "029480") == F) %>%
  dplyr::filter(str_detect(gene_ID, "030870") == F) %>%
  dplyr::filter(str_detect(gene_ID, "021170") == F) %>% 
  dplyr::filter(str_detect(gene_ID, "028970") == F) %>% 
  dplyr::filter(str_detect(gene_ID, "004670") == F)
  

Marker_dotplot <- DotPlot(CRO_RNA, 
                          features = unique(leaf_markers2$gene_ID)) 

Marker_dotplot_nice_df <- Marker_dotplot$data %>% 
  mutate(gene_ID = row.names(.)) %>% 
  mutate(gene_ID = str_sub(gene_ID, end = 15)) %>% 
  full_join(leaf_markers2 %>% 
              distinct(gene_ID, .keep_all = T),
             by = "gene_ID") %>% 
   mutate(cell_type2 = case_when(
    str_detect(cell_type, "meso") ~ "M",
    str_detect(cell_type, "cam|x|phl|com") ~ "V",
    str_detect(cell_type, "epi") ~ "E",
    str_detect(cell_type, "guard") ~ "GC",
    str_detect(cell_type, "IPAP") ~ "IP",
    str_detect(cell_type, "idioblast")  ~ "I"
  )) %>% 
  dplyr::filter(str_detect(cell_type, "shea") == F) %>% 
  mutate(cell_type2 = factor(cell_type2, levels = c(
    "M", "E", "GC", "V", "IP", "I"
  ))) %>% 
  mutate(cell_type3 = as.numeric(cell_type2)) %>% 
  mutate(gene_ID = reorder(gene_ID, -cell_type3))  %>% 
  mutate(id = factor(id, levels = c(
    "0", "1", "3", "7", 
    "2","5", 
    "12",
    "11", "6", "10", "14", "15",
    "13", "8",
    "4", "9"
  ))) %>% 
  dplyr::filter(is.na(avg.exp.scaled) == F)


Marker_dot <- Marker_dotplot_nice_df %>% 
  ggplot(aes(y = gene_ID, x = id)) +
  geom_point(aes(size = pct.exp, fill = avg.exp.scaled), shape = 21, color = "white") +
  scale_fill_viridis(option = "A", begin = 0, end = 0.9) +
  guides(size = guide_legend(override.aes = list(fill = "grey40"), nrow = 2)) +
  labs(x = "Cluster",
       y = NULL, 
       fill = "Avg. Exp.",
       size = "% Exp.") +
  theme_classic() +
  theme(text = element_text(size = 14, color = "black"),
        axis.text = element_text(color = "black"),
        legend.position = "top",
        legend.box = "horizontal",
        legend.key.height = unit(0.7, "lines"))

Marker_dot
```

* Mesophyll: 0, 1, 3, 7
* Epi: 2, 5
* Guard: 12,
* Vasculature: 11, 6, 10, 14, 15 
* Un: 4, 9 

* IPAP: 13
* Idioblast: 8 

```{r}
leaf_cell_type_assignment <- data.frame(
  cluster = c(0:15)
) %>% 
  mutate(cell_type = case_when(
    cluster == 0 | 
      cluster == 1 | 
      cluster == 3 | 
      cluster ==7  ~ "Mesophyll",
    cluster == 2 | 
      cluster == 5 ~ "Epidermis",
    cluster == 12 ~ "Guard cells",
    cluster == 11 |
      cluster == 6 |
      cluster == 10 |
      cluster == 14 |
      cluster == 15  ~ "Vasculature",
    cluster == 13 ~ "IPAP",
      cluster == 8 ~ "Idioblast",
    T ~ "Unassigned"
  )) %>% 
  mutate(cell_type = factor(cell_type, 
                            levels = c(
                              "Mesophyll", "Epidermis", "Guard cells",
                              "Vasculature", 
                              "IPAP", "Idioblast", "Unassigned"
                            ))) %>% 
  mutate(cluster = factor(cluster, levels = c(
    "0", "1", "3", "7", 
    "2","5", 
    "12",
    "11", "6", "10", "14", "15",
    "13", "8",
    "4", "9"
  )))
```


```{r}
Leaf_cell_type_strip <- leaf_cell_type_assignment %>% 
  ggplot(aes(x = cluster, y = "" )) +
  geom_tile(aes(fill = cell_type)) +
  scale_fill_manual(values = c(brewer.pal(6, "Accent"), "grey80")) +
  guides(fill = guide_legend(nrow = 2)) +
  labs(fill = "Cell type") +
  theme_void() +
  theme(
    legend.position = "bottom" ,
    text = element_text(size = 14)
  )

# Leaf_cell_type_strip

Leaf_cell_type_mkr_text <- Marker_dotplot_nice_df %>% 
  dplyr::filter(id == 0) %>% 
  ggplot(aes(x = "", y = gene_ID)) +
  geom_text(aes(label = Symbol)) +
  theme_void()

blank <- data.frame(
  x = 0,
  y = 0
) %>% 
  ggplot(aes(x, y)) +
  theme_void()

Marker_dot_strip <- wrap_plots(Marker_dot, Leaf_cell_type_mkr_text,
          Leaf_cell_type_strip, blank,
          heights = c(1, 0.03), 
          widths = c(1, 0.2), nrow = 2, ncol = 2)

Marker_dot_strip
ggsave("../Results/R_output/Cell_type_assignment_plot.svg", height = 6, width = 7.5, bg = "white")
ggsave("../Results/R_output/Cell_type_assignment_plot.png", height = 6, width = 7.5, bg = "white")
```

```{r}
MIA_heatmap_color_strip <- wrap_plots(MIA_heatmap +
                                        labs(tag = "C"), Leaf_cell_type_strip,
           nrow = 2, heights = c(1, 0.02))

MIA_heatmap_color_strip

ggsave("../Results/R_output/MIA_at_cluster.svg", height = 7.5, width = 5, bg = "white")
ggsave("../Results/R_output/MIA_at_cluster.png", height = 7.5, width = 5, bg = "white")
```

# UMAP with cell type info
```{r}
levels(CRO_RNA) <- c(
   "0", "1", "3", "7", 
    "2","5", 
    "12",
    "11", "6", "10", "14", "15",
    "13", "8",
    "4", "9"
  )

UMAP2 <- DimPlot(CRO_RNA, reduction = "umap", 
                 label = T, label.size = 5, repel = T) + 
  theme_void() +
  theme(
    text = element_text(size = 14, color = "black"),
    legend.position = "none"
  )

UMAP2
```
```{r}
leaf_assignment_graph <- leaf_cell_type_assignment %>% 
  ggplot(aes(x = cluster, y = "")) +
  facet_wrap(~ cell_type, scales = "free", ncol = 2) +
  geom_point(size = 5, aes(color = cluster)) +
  geom_text(aes(label = cluster)) +
  labs(x = NULL,
       y = NULL) +
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.grid = element_blank(),
    text = element_text(size = 14, color = "black"),
    axis.text.x = element_blank(),
    panel.border = element_rect(color = "black", fill = NA)
    )

RNA_umap <- wrap_plots(UMAP2 +
                         labs(tag = "B"), leaf_assignment_graph, 
           nrow = 2 , heights = c(1, 0.7))

RNA_umap
ggsave("../Results/R_output/UMAP_2.svg", height = 6, width = 3, bg = "white")
ggsave("../Results/R_output/UMAP_2.png", height = 6, width = 3, bg = "white")
```

# Fraction of cell types of interest 
```{r}
CRO_RNA@meta.data <- CRO_RNA@meta.data %>% 
  mutate(cell_type = case_when(
    seurat_clusters == 0 | 
      seurat_clusters == 1 | 
      seurat_clusters == 3 | 
      seurat_clusters ==7  ~ "Mesophyll",
    seurat_clusters == 2 | 
      seurat_clusters == 5 ~ "Epidermis",
    seurat_clusters == 12 ~ "Guard cells",
    seurat_clusters == 11 |
      seurat_clusters == 6 |
      seurat_clusters == 10 |
      seurat_clusters == 14 |
      seurat_clusters == 15  ~ "Vasculature",
    seurat_clusters == 13 ~ "IPAP",
      seurat_clusters == 8 ~ "Idioblast",
    T ~ "Unassigned"
  )) 


CRO_RNA@meta.data %>% 
  group_by(cell_type) %>% 
  dplyr::count() %>% 
  mutate(total = nrow(CRO_RNA@meta.data)) %>% 
  mutate(percentage = n/total) %>% 
  ungroup()

CRO_multiome@meta.data %>% 
  group_by(cell_type) %>% 
  dplyr::count() %>% 
  mutate(total = nrow(CRO_multiome@meta.data)) %>% 
  mutate(percentage = n/total) %>% 
  ungroup()
```
```{r}
RNA_by_rep <- CRO_RNA@meta.data %>% 
  group_by(cell_type, run) %>% 
  dplyr::count() %>% 
  inner_join(
    CRO_RNA@meta.data %>% 
  group_by(run) %>% 
  dplyr::count() %>% 
  ungroup() %>% 
  dplyr::rename(total = n), 
  by = "run"
  ) %>% 
  mutate(percentage = n/total * 100) %>% 
  ungroup() %>% 
  mutate(cell_type = factor(cell_type, 
                            levels = c(
                              "Mesophyll", "Epidermis", "Guard cells",
                              "Vasculature", 
                              "IPAP", "Idioblast", "Unassigned"
                            ))) %>% 
  mutate(cell_type = fct_rev(cell_type)) %>% 
  mutate(Run = case_when(
    run == "AZ" ~ "Rep1",
    run == "BB" ~ "Rep2",
    run == "BD" ~ "Rep3"
  )) %>% 
  ggplot(aes(x = Run, y = percentage)) +
  geom_bar(stat = "identity", aes(fill = cell_type)) +
  scale_fill_manual(values = rev(c(brewer.pal(6, "Accent"), "grey80"))) +
  labs(x = "", 
       y = "% cells by cell type",
       fill = "cell type") +
  theme_classic() +
  theme(legend.position = "none") +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, color = "black")) +
  coord_flip() 
```

```{r}
multiome_proportions <- CRO_multiome@meta.data %>% 
  group_by(cell_type) %>% 
  dplyr::count() %>% 
  mutate(total = nrow(CRO_multiome@meta.data)) %>% 
  mutate(percentage = n/total) %>% 
  ungroup() %>% 
  mutate(cell_type = factor(cell_type, 
                            levels = c(
                              "Mesophyll", "Epidermis", "Guard cells",
                              "Vasculature", 
                              "IPAP", "Idioblast", "Unassigned"
                            ))) %>% 
  mutate(cell_type = fct_rev(cell_type)) %>% 
  ggplot(aes(x = "Multiome", y = percentage)) +
  geom_bar(stat = "identity", aes(fill = cell_type)) +
  scale_fill_manual(values = rev(c(brewer.pal(6, "Accent"), "grey80"))) +
  labs(x = "", 
       y = "% cells by cell type",
       fill = "cell type") +
  theme_classic() +
  theme(legend.position = "none") +
  theme(axis.text.y = element_text(angle = 0, hjust = 1, color = "black")) +
  coord_flip() 
```


```{r}
wrap_plots(RNA_by_rep, multiome_proportions, nrow = 2, 
           heights = c(1, 0.3))

ggsave("../Results/R_output/Cell_proportions.svg", height = 2.3, width = 6.5, bg = "white")
ggsave("../Results/R_output/Cell_proportions.png", height = 2.3, width = 6.5, bg = "white")
```

```{r}
saveRDS(CRO_RNA, "../Results/R_output/CRO_RNA.Rds")
```

# Output figures 
```{r}
CRO_RNA <- readRDS("../Results/R_output/CRO_RNA.Rds")
CRO_multiome <- readRDS("../Results/R_output/CRO_multiome.Rds")
DefaultAssay(CRO_multiome) <- "RNA"
```

## Main 1 
```{r}
wrap_plots(
  combined_UMAPs, 
  RNA_umap,
  MIA_heatmap_color_strip, 
  design = c("AA
              BC"),
  heights = c(0.3, 1),
  widths = c(0.9, 1)
  )  

ggsave("../../Multiome_figures/Figure1.svg", height = 10, width = 8, bg = "white")
ggsave("../../Multiome_figures/Figure1.png", height = 10, width = 8, bg = "white")
```

## Sups 
```{r}
# wrap_plots(
#   UMAP_by_rep +
#     labs(tag = "A"),
#   Marker_dot_strip,
#   cell_proportions +
#     labs(tag = "C"),
#   nrow = 3, 
#   heights = c(0.35, 1, 0.15)
# )
# 
# ggsave("../../Multiome_figures/Sub_scRNA_2.svg", height = 10.5, width = 8, bg = "white")
# ggsave("../../Multiome_figures/Sub_scRNA_2.png", height = 10.5, width = 8, bg = "white")
```

```{r}
wrap_plots(
  AZ_QC,
  BB_QC,
  BD_QC,
  ncol = 1
) &
  theme(title = element_text(size = 10))

ggsave("../../Multiome_figures/Sub_scRNA_1.svg", height = 10.5, width = 6, bg = "white")
ggsave("../../Multiome_figures/Sub_scRNA_1.png", height = 10.5, width = 6, bg = "white")
```
## TFs 
```{r}
TFs <- readRDS("../Data/MIA_genes_info.Rds")
TFs <- TFs %>% 
  filter(is.na(tag) == F) %>% 
  filter(segment == "TF")

TFs
```

```{r}
TF_of_focus <- TFs %>% 
  filter(str_detect(tag, "ZCT") == F) %>% 
  mutate(LocusID = str_replace(cro_v3_gene, "_", "-")) %>% 
  inner_join(repr_loci, by = "LocusID") %>% 
  add_row(gene_ID = "CRO-03G006290.2", tag = "GATA1")

TF_of_focus
```
```{r}
TF_dot_data <- DotPlot(CRO_RNA, features = TF_of_focus$gene_ID)$data
head(TF_dot_data)
```
```{r}
TF_heatmap <- TF_dot_data%>% 
  mutate(internal.tag = row.names(.)) %>% 
  arrange(internal.tag) %>% 
  mutate(gene_ID = case_when(
    is.na(features.plot) ~ str_remove_all(internal.tag, "rna_"),
    T ~ features.plot %>% as.character()
  )) %>% 
  inner_join(TF_of_focus, by = "gene_ID") %>%
  mutate(id = factor(id, levels = c(
    "0", "1", "3", "7", 
    "2","5", 
    "12",
    "11", "6", "10", "14", "15",
    "13", "8",
    "4", "9"
  ))) %>% 
  ggplot(aes(y = tag, x = id)) +
  geom_point(aes(size = pct.exp, fill = avg.exp.scaled), shape = 21, color = "white") +
  scale_fill_viridis(option = "A", begin = 0, end = 0.9) +
  scale_size(breaks = c(10, 25, 50)) +
  guides(size = guide_legend(override.aes = list(fill = "grey40"))) +
  labs(x = "Cluster",
       y = NULL, 
       fill = "Avg.\nExp.",
       size = "% Exp.") +
  theme_classic() +
  theme(text = element_text(size = 14, color = "black"),
        axis.text = element_text(color = "black"),
        axis.text.y.left = element_text(hjust = 0.5),
        legend.key.width = unit(0.8, "lines"))

wrap_plots(TF_heatmap, Leaf_cell_type_strip,
           heights = c(1, 0.05), nrow = 2)

ggsave("../Results/R_output/TF_at_cluster.svg", height = 4.5, width = 6, bg = "white")
ggsave("../Results/R_output/TF_at_cluster.png", height = 4.5, width = 6, bg = "white")
```

## Supp. table for marker genes 
```{r}
leaf_markers3 <- leaf_markers %>% 
  dplyr::select(cell_type, Symbol, TAIR, gene_ID, cro.v3) %>% 
  #rbind(cro_special) %>% 
  dplyr::filter(str_detect(gene_ID, "007040") == F) %>% 
  dplyr::filter(str_detect(gene_ID, "004679") == F) %>%
  dplyr::filter(str_detect(gene_ID, "023500") == F) %>%
  dplyr::filter(str_detect(gene_ID, "008230") == F) %>%
  dplyr::filter(str_detect(gene_ID, "021030") == F) %>%
  dplyr::filter(str_detect(gene_ID, "029480") == F) %>%
  dplyr::filter(str_detect(gene_ID, "030870") == F) %>%
  dplyr::filter(str_detect(gene_ID, "021170") == F) %>% 
  dplyr::filter(str_detect(gene_ID, "028970") == F) %>% 
  dplyr::filter(str_detect(gene_ID, "004670") == F) %>% 
  filter(gene_ID %in% Marker_dotplot_nice_df$gene_ID) %>% 
  distinct(gene_ID, .keep_all = T) %>% 
  dplyr::select(-gene_ID)
  

leaf_markers3
```

```{r}
write_excel_csv(leaf_markers3, "../Results/R_output/leaf_marker3.csv")
```
 

