---
title: "Petal_pip_new"
author: "Chenxin Li"
date: "2024-05-14"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Packages
```{r}
library(tidyverse)
library(Seurat)

library(readxl)
library(RColorBrewer)
library(viridis)
library(rcartocolor)
library(patchwork)
```

# Data
```{r}
BZ_raw <- Read10X("../Results/cro_bz_em_emptydrops_combined/")
ca_raw <- Read10X("../Results/cro_ca_em_emptydrops_combined/")
```

# Object 
```{r}
BZ <- CreateSeuratObject(counts = BZ_raw, min.cells = 3, min.features = 200,
                         project = "BZ")

CA <- CreateSeuratObject(counts = ca_raw, min.cells = 3, min.features = 200,
                         project = "CA")
```

```{r}
BZ
CA
```
# Filtering
```{r}
blank <- data.frame(
  x = 1:10,
  y = 1:10
) %>% 
  ggplot(aes(x = x, y = y)) +
  theme_void()
```

## BZ
```{r}
BZ_scatter <- BZ@meta.data %>% 
  ggplot(aes(x = nCount_RNA, y = nFeature_RNA)) +
  geom_point(alpha = 0.5) +
  geom_vline(xintercept = 500, color = "red2") +
  geom_vline(xintercept = 30000, color = "red2") +
  geom_hline(yintercept = 300, color = "blue1") +
  geom_hline(yintercept = 10000, color = "blue1") +
  scale_x_log10() +
  scale_y_log10() +
  theme_classic()

BZ_hist1 <- BZ@meta.data %>% 
  ggplot(aes(x = nCount_RNA)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 500, color = "red2") +
  geom_vline(xintercept = 30000, color = "red2") +
  scale_x_log10() +
  theme_classic()

BZ_hist2 <- BZ@meta.data %>% 
  ggplot(aes(x = nFeature_RNA)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 300, color = "blue1") +
  geom_vline(xintercept = 10000, color = "blue1") +
  scale_x_log10() +
  theme_classic() +
  coord_flip()

wrap_plots(
  BZ_hist1, blank,
  BZ_scatter, BZ_hist2,
  nrow = 2, ncol = 2, 
  widths = c(1, 0.2), 
  heights = c(0.2, 1)
)

ggsave("../Results/R_output/BZ_filtering.svg", height = 4, width = 5, bg = "white")
ggsave("../Results/R_output/BZ_filtering.png", height = 4, width = 5, bg = "white")
```

## CA
```{r}
CA_scatter <- CA@meta.data %>% 
  ggplot(aes(x = nCount_RNA, y = nFeature_RNA)) +
  geom_point(alpha = 0.5) +
  geom_vline(xintercept = 500, color = "red2") +
  geom_vline(xintercept = 30000, color = "red2") +
  geom_hline(yintercept = 300, color = "blue1") +
  geom_hline(yintercept = 10000, color = "blue1") +
  scale_x_log10() +
  scale_y_log10() +
  theme_classic()

CA_hist1 <- CA@meta.data %>% 
  ggplot(aes(x = nCount_RNA)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 500, color = "red2") +
  geom_vline(xintercept = 30000, color = "red2") +
  scale_x_log10() +
  theme_classic()

CA_hist2 <- CA@meta.data %>% 
  ggplot(aes(x = nFeature_RNA)) +
  geom_histogram(bins = 100) +
  geom_vline(xintercept = 300, color = "blue1") +
  geom_vline(xintercept = 10000, color = "blue1") +
  scale_x_log10() +
  theme_classic() +
  coord_flip()

wrap_plots(
  CA_hist1, blank,
  CA_scatter, CA_hist2,
  nrow = 2, ncol = 2, 
  widths = c(1, 0.2), 
  heights = c(0.2, 1)
)

ggsave("../Results/R_output/CA_filtering.svg", height = 4, width = 5, bg = "white")
ggsave("../Results/R_output/CA_filtering.png", height = 4, width = 5, bg = "white")
```
```{r}
BZ_sub <- subset(BZ, subset = nFeature_RNA > 300 &
                   nCount_RNA > 500)

BZ_sub
```

```{r}
CA_sub <- subset(CA, subset = nFeature_RNA > 300 &
                   nCount_RNA > 500)

CA_sub
```

```{r}
lib_summary <- data.frame(
  ID = c("BZ", "CA"),
  cells = c(ncol(BZ_sub), 
            ncol(CA_sub)),
  median_UMI = c(summary(BZ_sub$nCount_RNA)[3],
                 summary(CA_sub$nCount_RNA)[3]),
  median_genes = c(summary(BZ_sub$nFeature_RNA)[3],
                   summary(CA_sub$nFeature_RNA)[3])
)

lib_summary

write_excel_csv(lib_summary, "../Results/R_output/lib_summary_petal_pip_new.csv")
```
# Normalize and variable features 
```{r}
BZ_sub <- NormalizeData(BZ_sub, normalization.method = "LogNormalize", scale.factor = 10000)
CA_sub <- NormalizeData(CA_sub, normalization.method = "LogNormalize", scale.factor = 10000)
```
```{r}
BZ_sub <- FindVariableFeatures(BZ_sub, selection.method = "vst", nfeatures = 3000)
CA_sub <- FindVariableFeatures(CA_sub, selection.method = "vst", nfeatures = 3000)
```
# Integrate
```{r}
RNA.list <- list(BZ_sub, CA_sub)
features <- SelectIntegrationFeatures(object.list = RNA.list, nfeatures = 3000)
anchors <- FindIntegrationAnchors(object.list = RNA.list, anchor.features = features) 
```
```{r}
CRO_petals <- IntegrateData(anchorset = anchors)
CRO_petals 
```
# Scale Integrated data 
```{r}
CRO_petals <- ScaleData(CRO_petals, verbose = F)
CRO_petals <- FindVariableFeatures(CRO_petals, selection.method = "vst", nfeatures = 3000)
  
CRO_petals@meta.data$run <- CRO_petals@meta.data$orig.ident

CRO_petals@meta.data %>% 
  group_by(run) %>% 
  dplyr::count()
```

# PCA 
```{r}
CRO_petals <- RunPCA(CRO_petals, verbose = FALSE, npcs = 50) # run PCA

DimPlot(CRO_petals, reduction = "pca", cols = brewer.pal(8, "Set2")[c(1:3)],
        split.by = "run", group.by = "run") +
theme(
  legend.position = "none"
)
```

# Clustering and UMAP 
```{r}
CRO_petals <- RunUMAP(CRO_petals, dims = 1:30, 
                     min.dist = 0.001, repulsion.strength = 1, 
                     n.neighbors = 30, spread = 1
                    ) 

CRO_petals <- FindNeighbors(CRO_petals, reduction = "pca", dims = 1:50)

CRO_petals <- FindClusters(CRO_petals, resolution = 0.5)



UMAP_RNA <- DimPlot(CRO_petals, reduction = "umap", 
                 label = T, label.size = 5, repel = T) + 
  theme_void() +
  theme(
    text = element_text(size = 14, color = "black", face = "bold"),
    legend.position = "none"
  )

UMAP_RNA

ggsave("../Results/R_output/UMAP_petal_pip2.svg", height = 3, width = 3, bg = "white")
ggsave("../Results/R_output/UMAP_petal_pip2.png", height = 3, width = 3, bg = "white")
```

## Check replicates 
```{r}
DimPlot(CRO_petals, reduction = "umap", 
        label = T, label.size = 5, repel = T, split.by = "run") +
  theme_void() +
  theme(
    text = element_text(size = 14, color = "black", face = "bold"),
    legend.position = "none",
    title = element_text(size = 10)
  ) +
  ggtitle("Grouped by replicates\n")  
  

ggsave("../Results/R_output/UMAP_RNA_by_rep_pip2.svg", height = 3, width = 8, bg = "white")
ggsave("../Results/R_output/UMAP_RNA_by_rep_pip2.png", height = 3, width = 8, bg = "white")
```
```{r}
saveRDS(CRO_petals, "../Results/R_output/petal_pip2.Rds")
CRO_petals <- readRDS("../Results/R_output/petal_pip2.Rds")
```

# Heat map for MIA genes 
## Annotation 
```{r}
repr_gene_models <- read_delim("../Data/cro_v3_anno/cro_v3.gene_models.repr.gff3",
                               delim = "\t", col_names = F)

repr_loci <- repr_gene_models %>% 
  dplyr::filter(X3 == "mRNA") %>% 
  dplyr::select(X1, X4, X5, X7, X9) %>% 
  separate(X9, c("ID", "Name"), sep = ";") %>% 
  dplyr::select(-Name) %>% 
  separate(ID, c("temp", "gene_ID"), sep = "=") %>% 
  dplyr::select(-temp) %>% 
  mutate(LocusID = str_sub(gene_ID, start = 1, end = 13)) %>% 
  mutate(gene_ID = str_replace(gene_ID, "_", "-")) %>% 
  mutate(LocusID = str_replace(LocusID, "_", "-")) %>% 
  dplyr::rename(
    Chr = X1,
    start = X4,
    end = X5,
    strand = X7
  )


head(repr_loci)
```
## Pull MIA gens 
```{r}
MIA_genes_info <- readRDS(file = "../Data/MIA_genes_info.Rds")

leaf_MIA_genes <- MIA_genes_info %>% 
  mutate(tag = case_when(
    str_detect(as.character(gene), "05G028810") ~ "7DLGT",
    str_detect(as.character(gene), "04G032090") ~ "GPPS(LSU)",
    str_detect(as.character(gene), "05G017180") ~ "DPAS", 
    T ~ tag
    )) %>% 
  dplyr::filter(is.na(tag) == F) %>% 
  mutate(order = case_when(
    str_detect(as.character(gene), "04G032090") ~  9,
    str_detect(as.character(gene), "05G028810") ~ 15, 
    str_detect(tag, "THAS2") ~ 40,
    T ~ order
  )) %>% 
  arrange(order) %>%
  mutate(order2 = case_when(
    str_detect(as.character(gene), "04G032090") ~  9,
    str_detect(as.character(gene), "05G028810") ~ 15, 
    str_detect(as.character(gene), "05G017180") ~ 31,
    order >= 31 ~ order + 1,
    T ~ order
  )) %>% 
  arrange(order2) %>% 
  mutate(segment = case_when(
    str_detect(as.character(gene), "04G032090") ~ "MEP",
    str_detect(as.character(gene), "05G028810") ~ "Iridoid", 
    T ~ segment
  )) %>% 
  dplyr::filter(is.na(order2) == F) %>% 
  dplyr::filter(str_detect(as.character(gene), "05G025350") == F) %>% 
  dplyr::filter(str_detect(as.character(gene), "05G028780") == F) %>% 
  dplyr::filter(str_detect(tag, "THAS1|THAS3|THAS4|HYS|ISP|GS2") == F) %>% 
  mutate(tag = case_when(
    str_detect(tag, "SGD$") ~ "SGD1",
    #str_detect(tag, "SGD2") ~ "SDG2",
    str_detect(tag, "GS$") ~ "GS1",
    str_detect(tag, "MATE") ~ "SLTr",
    T ~ tag
  )) %>% 
  mutate(gene = str_replace_all(gene, "_", "-")) %>% 
  left_join(repr_loci %>% 
              dplyr::select(-Chr, -start, -end, -strand), by = c("gene"="LocusID"))

head(leaf_MIA_genes)
write_excel_csv(leaf_MIA_genes, "../Results/R_output/leaf_MIA_genes.csv")
```
```{r}
MIA_genes_info %>% 
  filter(is.na(tag) == F)
```


## Plot
```{r}
DefaultAssay(CRO_petals) <- "RNA"

Dotplot_for_makers <- leaf_MIA_genes %>% 
  filter(order > 36) %>% 
  select(gene, tag, order2) %>% 
  rbind(
    data.frame(
      gene = "CRO-04G032540",
      tag = "KCS",
      order2 = -10
    )
  ) %>% 
  rbind(
    data.frame(
      gene = "CRO-04G030450",
      tag = "F3H",
      order2 = -3
    )
  ) %>%
  rbind(
    data.frame(
      gene = "CRO-04G031350",
      tag = "DFR",
      order2 = -2
    )
  ) %>% 
  rbind(
    data.frame(
      gene = "CRO-07G003740",
      tag = "ANS",
      order2 = -1
    )
  ) 
Dotplot_for_makers
MIA_at_cluster <- DotPlot(CRO_petals, features = c(rev(Dotplot_for_makers$gene)))
```
```{r}
MIA_heatmap <- MIA_at_cluster$data %>% 
  mutate(internal.tag = row.names(.)) %>% 
  arrange(internal.tag) %>% 
  mutate(gene = case_when(
    is.na(features.plot) ~ str_remove_all(internal.tag, "rna_"),
    T ~ features.plot %>% as.character()
  )) %>% 
  mutate(gene = str_sub(gene, start = 1, end = 13)) %>% 
  inner_join(Dotplot_for_makers, by = "gene") %>%
  mutate(gene = reorder(gene, -order2)) %>%
  mutate(tag = reorder(tag, -order2)) %>% 
  mutate(id = factor(id, levels = c(
 "0","2", "13", 
  "5", "6", 
  "10",  
  "1", "4", "8",
  "3", "7", "9", "11" ,"12", "14" 
  ))) %>%
  ggplot(aes(y = tag, x = id)) +
  geom_point(aes(size = pct.exp, fill = avg.exp.scaled), shape = 21, color = "white") +
  scale_fill_viridis(option = "A", begin = 0, end = 0.9) +
  guides(size = guide_legend(override.aes = list(fill = "grey40"))) +
  labs(x = "Cluster",
       y = NULL, 
       fill = "Avg.\nExp.",
       size = "% Exp.") +
  theme_classic() +
  theme(text = element_text(size = 12, color = "black"),
        axis.text = element_text(color = "black"),
        axis.text.y.left = element_text(hjust = 0.5, face = "italic"),
        legend.key.width = unit(0.75, "lines"))

MIA_heatmap
```

# Assign cell types

* Epidermis: 0, 2, 13 
* Parenchyma: 1, 3,  4, 8 
* Pigment: 5, 6
* Idioblast: 10 
* Unassigned: 3, 7, 9, 11, 12, 14  

```{r}
petal_cell_type_assignment <- data.frame(
  cluster = c(0:14)
) %>% 
  mutate(cell_type = case_when(
    cluster == 0 | 
      cluster == 2 | 
      cluster == 13  ~ "Epidermis",
    cluster == 1 | 
      cluster == 3 |
      cluster == 4 |
      cluster == 8 ~ "Parenchyma",
    cluster == 5 | 
      cluster == 6 ~ "Pigment", 
    cluster == 10 ~ "Idioblast", 
    T ~ "Unassigned"
  )) %>% 
  mutate(cell_type = factor(cell_type, 
                            levels = c(
                              "Epidermis", "Pigment", "Idioblast", "Parenchyma", "Unassigned"
                            ))) %>% 
  mutate(cluster = factor(cluster, levels = c(
  "0","2", "13", 
  "5", "6", 
  "10",  
  "1", "4", "8",
  "3", "7", "9", "11" ,"12", "14"
  )))
```

## cell type color strip 
```{r}
Petal_cell_type_strip <- petal_cell_type_assignment %>% 
  ggplot(aes(x = cluster, y = "" )) +
  geom_tile(aes(fill = cell_type)) +
   scale_fill_manual(values = c(
                               brewer.pal(8, "Accent")[2],
                               brewer.pal(8, "Pastel1")[1],
                               #brewer.pal(8, "Set2")[6],
                               brewer.pal(8, "Accent")[6],
                               "grey90",
                               "grey80")) +
  guides(fill = guide_legend(nrow = 3)) +
  labs(fill = "Cell type") +
  theme_void() +
  theme(
    legend.position = "bottom" ,
    text = element_text(size = 12)
  )

wrap_plots(MIA_heatmap, Petal_cell_type_strip, nrow = 2, heights = c(1, 0.05))


ggsave("../Results/R_output/late_MIA_dotplot.svg", height = 5, width = 4.5)
ggsave("../Results/R_output/late_MIA_dotplot.png", height = 5, width = 4.5)
```


# Fraction of idioblasts?
```{r}
CRO_petals@meta.data %>% 
  filter(seurat_clusters==10) %>% 
  nrow()

nrow(CRO_petals@meta.data)

86 / 5924 * 100
```

1.45% 

# Inspect de novo marker genes 
```{r}
de_novo_markers <- FindAllMarkers(CRO_petals, only.pos = T, min.pct = 0.25, logfc.threshold = 0.25, 
                                  verbose = F)

head(de_novo_markers)
```
```{r}
annotation <- read_delim("../Data/cro_v3_anno/cro_v3.functional_annotation.txt",
                         delim = "\t", col_names = F)

head(annotation)
```

```{r}
de_novo_markers %>% 
  left_join(annotation %>% 
              mutate(X1 = str_remove_all(X1, "\\.\\d+")) %>% 
              distinct(X1, .keep_all = T) %>% 
              mutate(gene = str_replace(X1, "_", "-")), by = "gene") %>% 
  filter(p_val_adj < 0.01) %>%
write_excel_csv("../Results/R_output/petal_de_novo_markers.csv")

```

## Visualize 
```{r}
selected_markers <- de_novo_markers %>% 
  filter(p_val_adj < 0.01) %>% 
  group_by(cluster) %>% 
  slice_max(order_by = avg_log2FC, n = 10) %>% 
  ungroup() 

selected_markers
```
```{r}
DotPlot(CRO_petals, features = unique(selected_markers$gene)) +
  coord_flip()+
  theme(axis.text.y = element_blank())
```

```{r}
petal_idioblas_markers <- de_novo_markers %>% 
  filter(cluster == 10) %>% 
  filter(p_val_adj < 0.05) %>%
  left_join(annotation %>% 
              mutate(X1 = str_remove_all(X1, "\\.\\d+")) %>% 
              distinct(X1, .keep_all = T) %>% 
              mutate(gene = str_replace(X1, "_", "-")), by = "gene") 
```

## Compare with leaf idioblast markers 
```{r}
CRO_RNA <- readRDS("../../CRO_multiome/Results/R_output/CRO_RNA.Rds")
DefaultAssay(CRO_RNA) <- "RNA"

leaf_de_novo_markers <- FindAllMarkers(CRO_RNA, only.pos = T, min.pct = 0.25, logfc.threshold = 0.25, 
                                  verbose = F)

head(leaf_de_novo_markers)
```
```{r}
leaf_idioblas_markers <- leaf_de_novo_markers %>% 
  filter(cluster == 8) %>% 
  filter(p_val_adj < 0.05) %>%
  left_join(annotation %>% 
              #mutate(X1 = str_remove_all(X1, "\\.\\d+")) %>% 
              #distinct(X1, .keep_all = T) %>% 
              mutate(gene = str_replace(X1, "_", "-")), by = "gene") %>% 
  mutate(gene2 = str_remove(gene, "\\.\\d+"))

head(leaf_idioblas_markers)

write_excel_csv(leaf_idioblas_markers, "../Results/leaf_idioblast_markers.csv")
```

```{r}
B_and_A <- petal_idioblas_markers %>% 
  filter(gene %in% str_remove(row.names(CRO_RNA),"\\.\\d+")) %>% 
  nrow() # 106

C_and_A <- leaf_idioblas_markers %>% 
  filter(gene2 %in% row.names(CRO_petals)) %>% 
  nrow() # 147 

U <- intersect(
  str_remove(row.names(CRO_RNA),"\\.\\d+"), 
  row.names(CRO_petals)
) %>% 
  length() # 16709

A <- intersect(
  petal_idioblas_markers$gene,
  leaf_idioblas_markers$gene2
) %>% 
  length()

B <- B_and_A - A
C <- C_and_A - A
D <- U - A - B - C
  
fisher.test(rbind(
  c(A, B),
  c(C, D)
))

A/B/(C/D)
```
```{r}
exp <- C_and_A / U
obs <- A / B_and_A

obs/exp

chisq.test(
  c(A, B_and_A), p = c(exp, 1 - exp)
)
```

```{r}
union(
  petal_idioblas_markers$gene,
  leaf_idioblas_markers$gene2
) %>% 
  length()

33/224
```



# Check TFs
```{r}
TF_dot <- DotPlot(CRO_petals, features = c(
 "CRO-03G000120", # WRKY
 "CRO-05G006800", # Myb1
 "CRO-04G033370", # Myb2
 "CRO-07G002170" # Myb3 - TT2a 
))

#TF_dot
```

```{r}
TF_dot_nice <- TF_dot$data %>% 
  mutate(gene = features.plot) %>% 
  mutate(tag = case_when(
    str_detect(gene, "03G000120") ~ "IDW1",
    str_detect(gene, "05G006800") ~ "IDM1",
    str_detect(gene, "04G033370") ~ "IDM2",
    str_detect(gene, "07G002170") ~ "IDM3"
  )) %>% 
   mutate(id = factor(id, levels = c(
  "0","2", "13", 
  "5", "6", 
  "10",  
  "1", "4", "8",
  "3", "7", "9", "11" ,"12", "14"
  ))) %>%
  ggplot(aes(y = tag, x = id)) +
  geom_point(aes(size = pct.exp, fill = avg.exp.scaled), shape = 21, color = "white") +
  scale_fill_viridis(option = "A", end = 0.9, breaks = c(0, 2)) +
  guides(size = guide_legend(override.aes = list(fill = "grey40"), 
                             order = 2, ncol = 1)) +
  labs(x = "Cluster",
       y = NULL, 
       fill = "Avg.\nExp.",
       size = "% Exp.") +
  theme_classic() +
  theme(
    text = element_text(size = 12, color = "black"),
    axis.text = element_text(color = "black"),
    axis.text.y = element_text(hjust = 0.5, face = "italic"),
    legend.position = "right",
    legend.box = "vertical", 
    legend.key.width = unit(0.8, "lines"),
    legend.title = element_text(hjust = 0.5),
    strip.text = element_blank() 
 )

wrap_plots(TF_dot_nice, Petal_cell_type_strip, nrow = 2, heights = c(1, 0.05))


ggsave("../Results/R_output/petal_TFs_pip2.svg", height = 4.8, width = 4.5, bg = "white")
ggsave("../Results/R_output/petal_TFs_pip2.png", height = 4.8, width = 4.5, bg = "white")
```

 

